---
title: "PA7201_CO2_emissions"
date: "20/11/2020"
output:
  pdf_document: default
  word_document: default
---
```{r, echo = FALSE, message = FALSE}
library(tinytex)
library(knitr)
library(tidyverse)
library(png)
source("C:/Users/44792/Desktop/Google Earth Engine/PA7201_Global_CO2_Emissions/plot_gas.R")
setwd("C:/Users/44792/Desktop/Google Earth Engine/PA7201_Global_CO2_Emissions/")
```


```{r, echo = FALSE, message = FALSE}
plot_gas("ee-chart_AUS.csv", 
         "CO column density", 
         "(mol/m^2)",
         "Australia")


#img_AUS <- readPNG()
```

Figure 1: Time series of CO column density (mol/m^2^) throughout 2020 in Australia (top)
Area averaged map of CO column density (mol/m^2^) throughout 2020 in Australia (bottom)

```{r, echo = FALSE, message = FALSE}
plot_gas("ee-chart_BR.csv", 
         "CO column density", 
         "(mol/m^2)",
         "Brazil")

#img_BR <- readPNG()
```

Figure 2: Time series of CO column density (mol/m^2^) throughout 2020 in Brazil (top)
Area averaged map of CO column density (mol/m^2^) throughout 2020 in Brazil (bottom)

```{r, echo = FALSE, message = FALSE}
plot_gas("ee-chart_DRC.csv", 
         "CO column density", 
         "(mol/m^2)",
         "DRC")

#img_DRC <- readPNG()
```

Figure 3: Time series of CO column density (mol/m^2^) throughout 2020 in Democratic Republic of Congo (DRC) (top)
Area averaged map of CO column density (mol/m^2^) throughout 2020 in DRC (bottom)

```{r, echo = FALSE, message = FALSE}
plot_gas("ee-chart_ID.csv", 
         "CO column density", 
         "(mol/m^2)",
         "Indonesia")

#img_ID <- readPNG()
```

Figure 4: Time series of CO column density (mol/m^2^) throughout 2020 in Indonesia (top)
Area averaged map of CO column density (mol/m^2^) throughout 2020 in Indonesia (bottom)

```{r, echo = FALSE, message = FALSE}
plot_gas("ee-chart_IN.csv", 
         "CO column density", 
         "(mol/m^2)",
         "India")

#img_IN <- readPNG()
```

Figure 5: Time series of CO column density (mol/m^2^) throughout 2020 in India (top)
Area averaged map of CO column density (mol/m^2^) throughout 2020 in India (bottom)

```{r, echo = FALSE, message = FALSE}
plot_gas("ee-chart_UG.csv", 
         "CO column density", 
         "(mol/m^2)",
         "India")

#img_UG <- readPNG()
```

Figure 6: Time series of CO column density (mol/m^2^) throughout 2020 in Uganda (top)
Area averaged map of CO column density (mol/m^2^) throughout 2020 in Uganda (bottom)

## Analysis

CO column density is shown in 6 regions around the globe. All regions display a large range in CO with some seasonal patterns visible. 

### Australia

Australia displays maximum CO in January (0.045 mol/m2) and October (0.039 mol/m2). January, 2020 brought a series of bushfires throughout Australia, leading to increased CO throughout the region. CO levels were stable from February to August, ranging from 0.018 - 0.024. Australia underwent a series of 'lockdowns' from March to October to halt the spread of COVID-19. Minor decreases in CO throughout this period may be caused by this action, but further analysis is needed to confirm that. Bushfire season for 2020-2021 begins again in September and this is apparent from the time series. Increased C) at this period is a clear signal of burning vegetation throughout the region     